---
title: "MSD QC Report (PDF)"
# Parameters expected by app.R (observeEvent -> quarto_render)
params:
  ab: ""                    # e.g., "G", "A", "M"
  plate: 0                  # e.g., 1..N
  data_path: ""             # RDS created by DataTableFunc(): list(controls, standards, qc, unknowns)
  user_id: ""
  analyzed_on: ""
  experiment_name: ""
  experiment_date: ""
  s1_raw_path: ""           # path to RDS saved from session$userData$s1_raw
  s2_raw_path: ""           # path to RDS saved from session$userData$s2_raw
  s3_processed_path: ""     # path to RDS for this plate only (G1/A3/etc.)
format:
  pdf:
    toc: true
    toc-depth: 2
    number-sections: false
header-includes:
  - \\usepackage{booktabs}
  - \\usepackage{longtable}
  - \\usepackage{array}
  - \\usepackage{multirow}
  - \\usepackage{wrapfig}
  - \\usepackage{float}
  - \\usepackage{colortbl}
  - \\usepackage{xcolor}
execute:
  echo: false
  warning: false
  message: false
---

```{r}
# ---- Setup -------------------------------------------------------------------
suppressPackageStartupMessages({
  library(dplyr)
  library(knitr)
  library(kableExtra)
  library(stringr)
})

# ---- Robust Helpers ----------------------------------------------------------

safe_readRDS <- function(path) {
  # Treat NULL/length-0/"" robustly
  if (is.null(path) || length(path) == 0) return(NULL)
  if (!is.character(path) || length(path) != 1 || !nzchar(path)) return(NULL)
  tryCatch(readRDS(path), error = function(e) NULL)
}

# Strictly scalar TRUE/FALSE for non-empty character
safe_nzchar <- function(x) {
  !is.null(x) && is.character(x) && length(x) == 1 && !is.na(x) && nzchar(x)
}

# Safe string with default
sstr <- function(x, default = "N/A") if (safe_nzchar(x)) x else default

# LaTeX table helper (no HTML)
kblx <- function(df, caption = NULL, digits = 3, align = "l", ...) {
  if (is.null(df) || !is.data.frame(df) || nrow(df) == 0) {
    return(knitr::asis_output("\\emph{Not available.}"))
  }
  knitr::kable(
    df, format = "latex", booktabs = TRUE, longtable = TRUE,
    caption = caption, digits = digits, align = align, ...
  ) %>%
    kableExtra::kable_styling(latex_options = c("hold_position"))
}

# Coerce numeric-ish
to_num <- function(x) {
  x <- ifelse(x %in% c(".", "NaN", "NA", "", "N/A"), NA, x)
  suppressWarnings(as.numeric(x))
}

# Badge-like text for PDF
badge_text <- function(text) {
  paste0("\\textbf{", text, "}")
}

# ---- Derive params safely ----------------------------------------------------
ab    <- if (safe_nzchar(params$ab)) params$ab else ""
plate <- if (!is.null(params$plate) && length(params$plate) == 1 && !is.na(params$plate)) params$plate else ""
plate_id <- paste0(ab, plate)

# Pretty Experiment Header values (robust fallbacks)
exp_name <- sstr(params$experiment_name, "N/A")
exp_user <- sstr(params$user_id,         "N/A")
exp_date <- sstr(params$experiment_date, "N/A")
run_time <- if (safe_nzchar(params$analyzed_on)) params$analyzed_on else format(Sys.time(), "%d%b%Y %I:%M %p")

# Paths (also printed later in debug)
path_s1 <- sstr(params$s1_raw_path, "")
path_s2 <- sstr(params$s2_raw_path, "")
path_s3 <- sstr(params$s3_processed_path, "")
path_data <- sstr(params$data_path, "")

# ---- Load inputs written by app.R -------------------------------------------
s1_raw <- safe_readRDS(path_s1)      # columns: Experiment, Information
s2_raw <- safe_readRDS(path_s2)      # columns: Reagent, Lot, ExpiryDate
s3_one <- safe_readRDS(path_s3)      # processed metadata frame for this plate
tables <- safe_readRDS(path_data)    # list(controls, standards, qc, unknowns)

# Normalize table names safely
table_names <- if (is.list(tables) && length(tables) > 0) tolower(names(tables)) else character(0)

# Convenience pulls (NULL if absent)
tbl_controls  <- if ("controls"  %in% table_names) tables[[which(table_names=="controls")]]  else NULL
tbl_standards <- if ("standards" %in% table_names) tables[[which(table_names=="standards")]] else NULL
tbl_unknowns  <- if ("unknowns"  %in% table_names) tables[[which(table_names=="unknowns")]]  else NULL
tbl_qc        <- if ("qc"        %in% table_names) tables[[which(table_names=="qc")]]        else NULL
```

# MSD QC Report — `r exp_name`

**Generated by:** `r exp_user`  
**Experiment Date:** `r exp_date`  
**Analyzed on:** `r run_time`  
**Plate:** `r plate_id`

---

## Plate-Level QC Summary

```{r, results='asis'}
if (is.data.frame(tbl_qc) && nrow(tbl_qc) > 0) {
  res <- dplyr::slice_head(tbl_qc, n = 1)
  qc_value <- toupper(as.character(if (!is.null(res$QC)) res$QC else ""))
  if (qc_value == "PASS") {
    knitr::asis_output(badge_text("PASS"))
  } else if (qc_value == "REPEAT") {
    knitr::asis_output(badge_text("REPEAT"))
  } else {
    knitr::asis_output(badge_text("N/A"))
  }
} else {
  knitr::asis_output(badge_text("N/A"))
}
```

> **Note:** This plate summary is computed from **Controls** and **Standards** according to the QC rules in `DataTable.R`.

---

## Experiment Metadata

```{r, results='asis'}
if (is.data.frame(s1_raw) && nrow(s1_raw) > 0) {
  if (all(c("Experiment","Information") %in% names(s1_raw))) {
    kblx(s1_raw, caption = "Section 1 — Experiment Metadata", align = "l")
  } else {
    kblx(as.data.frame(s1_raw), caption = "Section 1 — Experiment Metadata (raw)")
  }
} else {
  knitr::asis_output("\\emph{Section 1 (Experiment Metadata) is not available.}")
}
```

---

## Reagents and Lot Information

```{r, results='asis'}
if (is.data.frame(s2_raw) && nrow(s2_raw) > 0) {
  show_cols <- intersect(c("Reagent","Lot","ExpiryDate"), names(s2_raw))
  kblx(s2_raw[, show_cols, drop = FALSE], caption = "Section 2 — Reagents & Lots")
} else {
  knitr::asis_output("\\emph{Section 2 (Reagents and Lots) is not available.}")
}
```

---

## Plate Metadata Summary

```{r, results='asis'}
if (is.data.frame(s3_one) && nrow(s3_one) > 0) {
  preferred <- c("Plate","Read Time","Barcode1","Barcode2","Barcode3","Serial #",
                 "Experiment Date","User ID","Plate Type")
  cols <- intersect(preferred, names(s3_one))
  if (length(cols) == 0) cols <- names(s3_one)
  kblx(s3_one[, cols, drop = FALSE], caption = paste0("Section 3 — Plate Metadata (", plate_id, ")"))
} else {
  knitr::asis_output("\\emph{Section 3 (Plate Metadata) is not available.}")
}
```

---

## Standards

```{r, results='asis'}
if (is.data.frame(tbl_standards) && nrow(tbl_standards) > 0) {
  df <- as.data.frame(tbl_standards)
  round_if <- function(df, nm, d) { if (nm %in% names(df)) df[[nm]] <- round(df[[nm]], d); df }
  df <- round_if(df, "PeRecMean", 2)
  df <- round_if(df, "Calc_Conc_CV", 2)
  df <- round_if(df, "FStatRSquare", 4)
  df <- round_if(df, "MeanLOQ", 2)
  df <- round_if(df, "CVLOQ", 2)
  kblx(df, caption = "Standards")
} else {
  knitr::asis_output("\\emph{No data in Standards table.}")
}
```

---

## Controls

```{r, results='asis'}
if (is.data.frame(tbl_controls) && nrow(tbl_controls) > 0) {
  df <- as.data.frame(tbl_controls)
  for (nm in c("Conc","PeRecMean","Calc_Conc_Mean","Calc_Conc_CV","MeanLOQ","CVLOQ")) {
    if (nm %in% names(df)) df[[nm]] <- suppressWarnings(round(as.numeric(df[[nm]]), 2))
  }
  kblx(df, caption = "Controls")
} else {
  knitr::asis_output("\\emph{No data in Controls table.}")
}
```

---

## Unknowns

```{r, results='asis'}
if (is.data.frame(tbl_unknowns) && nrow(tbl_unknowns) > 0) {
  df <- as.data.frame(tbl_unknowns)
  preferred <- c(
    "Sample","Assay","Plate_Name",
    "CalcConc_InRange","Meaninrange","CVinrange",
    "MeanLOQ","CVLOQ","Thres","ThresCall",
    "QC_CalcConc","QC_InRange","QC_LOQ","QC_DilutionCheck","Sample_QC"
  )
  keep <- intersect(preferred, names(df))
  if (length(keep) > 0) df <- df[, keep, drop = FALSE]
  num_like <- intersect(c("CalcConc_InRange","Meaninrange","CVinrange","MeanLOQ","CVLOQ","Thres"), names(df))
  for (nm in num_like) df[[nm]] <- suppressWarnings(round(as.numeric(df[[nm]]), 3))
  kblx(df, caption = "Unknowns")
} else {
  knitr::asis_output("\\emph{No data in Unknowns table.}")
}
```

---

## Additional QC Tables (If Present)

```{r, results='asis'}
other <- setdiff(table_names, c("controls","standards","unknowns","qc"))
if (length(other) == 0) {
  knitr::asis_output("\\emph{No additional QC tables found.}")
} else {
  for (nm in other) {
    ot <- tables[[which(table_names == nm)]]
    cat("### ", tools::toTitleCase(nm), "\n\n", sep = "")
    if (is.data.frame(ot) && nrow(ot) > 0) {
      print(kblx(as.data.frame(ot), caption = paste0(tools::toTitleCase(nm), " Table")))
    } else {
      knitr::asis_output(paste0("\\emph{No data in ", nm, " table.}"))
    }
  }
}
```

---

## Debug Snapshot (Optional)

```{r, results='asis'}
cat("\\begin{quote}\\small\n")
cat("Plate ID: ", plate_id, "\n", sep = "")
cat("s1_raw_path: ", path_s1, "\n", sep = "")
cat("s2_raw_path: ", path_s2, "\n", sep = "")
cat("s3_processed_path: ", path_s3, "\n", sep = "")
cat("data_path: ", path_data, "\n", sep = "")
cat("Tables loaded: ", paste0(table_names, collapse = ", "), "\n", sep = "")
cat("\\end{quote}\n")
```
